---
title: "pstat131_project"
author: "Michael Rhine-6342398"
date: "12/16/2020"
output: html_document
---

**PSTAT 131 FINAL PROJECT: 2020 ELELECTION ANALYSIS**

**Professor Guo Yu**
**Michael Rhine 6342398**
**December 17, 2020**


# Introduction and Data {.tabset}

## Abstract
  There is no better time than now to work on the 2020 United States presidential election data! Despite that the 2016 presidential election came as a big surprise to many, Biden’s victory in the 2020 presidential election had been widely predicted before Nov 3rd.

  Predicting voter behavior is complicated for many reasons despite the tremendous effort in collecting, cleaning, analyzing, and understanding many available datasets. For this project, I will use cencus information and build mutliple classification models predict county-level winners in the 2020 election. Principal Component Analysis and Clustering Methods were also performed. However, at this time, PCA results have not been used in model development, and a proper analysis of our clusterings have not been done. This will be addressed at a later time. Finally, I developed several regularized regression models to estimate vote totals in each county and see who would have won the 2020 election. I also applied my knowledge of regression analysis to develop a custom model.
  
  Important variables in many of these models were found to be: transit, TotalPop, White, Minority(not a perfect colinearity), VotingAgeCitizen, FamilyWork, and Service. 
  
## Data
  Two datasets will be used in this project and were kindly provided by the course instructor, Professor Guo Yu. 
  
  Our first dataset is the 2020 Presidential Election Data. This dataset will mainly serve as the data of our Response variable. The dimensions of the dataset are 31167 observations with 5 columns: 
  state - The state the observation was in. There were 51 unique observations, the 50 states including Washington D.C.
  county - The county the observation was in.
  candidate - The presidential candidate. This will be our response for our classfication models.
  Party - The party of the candidate
  Votes - How many votes in were assigned to the candidate by county. This will be used to determine our response for classification, as well as serve as our response for regression.
  
  Next is second dataset, 2020 Census Data. This dataset will mostly contain our predictor information, as well as meta-information that will be used to merge the two datasets.
  A header a the Census Data is listed below. Most of the data is in percentages. Those that aren't, Men and Women, will be, besides TotalPop and qualitive data.

# Election and Census Data
```{r setup, include=FALSE, echo=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(maps)
library(ROCR)
library(ggridges)
library(dendextend)
library(e1071)
library(ISLR)
library(tree)
library(maptree)
library(glmnet)
library(class)
library(FNN)
library(randomForest)
library(gbm)
```
```{r, echo=FALSE, include=FALSE, message = F}
## read data and convert candidate names and party names from string to factor
election.raw <- read_csv("candidates_county.csv", col_names = TRUE) %>% 
  mutate(candidate = as.factor(candidate), party = as.factor(party))

## remove the word "County" from the county names
words.to.remove = c("County")
remove.words <- function(str, words.to.remove){
  sapply(str, function(str){
    x <- unlist(strsplit(str, " "))
    x <- x[!x %in% words.to.remove]
    return(paste(x, collapse = " "))
  }, simplify = "array", USE.NAMES = FALSE)
}
election.raw$county <- remove.words(election.raw$county, words.to.remove)

## read census data
census <- read_csv("census_county.csv")
census$County <- remove.words(census$County, words.to.remove)
```
```{r, echo=FALSE, include=FALSE}
#Question 1

#Dimensions
dim(election.raw)
#Total # of distinct values
length(unique(election.raw$state))

#Question 2

#Dimensions
dim(census)
#Check for missing Data
sum(is.na(census))
#Total # of distinct values for county & State
paste("Number of distinct census counties: ", length(unique(census$County)) )
paste("Number of distinct election.raw counties: ", length(unique(election.raw$county)) )
```
```{r, echo=F, message = F}
head(census)
```  
  The dimensions for election.raw is 31,167 rows by 5 columns. The 'states' parameter has 51 unique values, meaning all states and the District of Columbia are contained in our data.
  
  Our cencus data has 3220 observations with 37 parameters. We should not that we have more unique counties in the election dataset than our census dataset 2825 unique counties vs 1955 census. This will need to be handled when merging datasets. There is one missing value in the census dataset. This doesn't need to be handled now, but I'd like to just get it out of the way considering we only have one value missing. The missing value is in childPoverty. Since it is only one observation, the average ChildPoverty value was computed and put in to replace it.
```{r,echo=F,include=F, message = F}
#Find missing data
apply(is.na(census),2,which)
```

```{r,echo=F, include = F, message = F}
#Replace missing values with mean value for ChildPoverty
new_val = mean(census$ChildPoverty, na.rm=TRUE)
census$ChildPoverty[is.na(census$ChildPoverty)] = new_val
#Double Checking for missing data
apply(is.na(census),2,which)
```

# Data Wrangling

Next we'll construct aggregated data sets from election.raw data:

```{r, echo=F, message = F}
#Question 3

#election.county
election.county <- election.raw
head(election.county)
```
election.county will be a direct copy of election.raw
```{r,echo=F, message = F}
#Question 3

#election.state
election.state <- election.raw %>% group_by(state,candidate) %>% summarise(votes=sum(votes))
head(election.state)
```
election.state will be an aggregate of votes per candidate by state.
```{r,echo=F, message = F}
#Question 3

#election.total
election.total <- election.state %>% group_by(candidate) %>% summarise(votes=sum(votes))
election.total <- election.total[order(election.total$votes, decreasing=T),]
head(election.total)
```
election.total will be total votes per candidate.

```{r, echo=F, message = F}
#Question 4

p <- ggplot(data=election.total, aes(x=reorder(candidate,-votes))) + 
  geom_bar(aes(y=log(votes)), stat='identity', fill='steelblue') + ylab("log(votes)")

p + theme(axis.text.x=element_text(color='black',size=7,angle=90), axis.title.x = element_blank())+ labs(title='Votes per candidate (in log)')

#remove scientific notation
options(scipen=999)
#Seperating election total into subsets by votes
el.tot_1 = election.total[1:3,]
el.tot_2 = election.total[3:12,]
el.tot_3 = election.total[13:38,]


el.tot_1.plot <- ggplot(data=el.tot_1, aes(x=reorder(candidate,-votes))) + 
  geom_bar(aes(y=votes), stat='identity', fill='steelblue') 

el.tot_2.plot <- ggplot(data=el.tot_2, aes(x=reorder(candidate,-votes))) + 
  geom_bar(aes(y=votes), stat='identity', fill='steelblue') 

el.tot_3.plot <- ggplot(data=el.tot_3, aes(x=reorder(candidate,-votes))) + 
  geom_bar(aes(y=votes), stat='identity', fill='steelblue')

grid.arrange(el.tot_1.plot +
  theme(axis.text.x=element_text(color='black',size=7,angle=90),axis.title.x=element_blank()) ,
             el.tot_2.plot +
  theme(axis.text.x=element_text(color='black',size=7,angle=90),axis.title.x=element_blank()) ,
             el.tot_3.plot  +
  theme(axis.text.x=element_text(color='black',size=7,angle=90),axis.title.x=element_blank()),
             layout_matrix = rbind(c(1,2,NA),
                                   c(3,3,NA)))
```

  We can take the number of dimensions for election.total to get the number of candidates voted on in the election. There were 38 overall. Notably, we can see Kanye West is the 8th column from the left. However, it would not be correct to say that the election was close among more than 2 candidates. Joe Biden and Donald Trump were the clear frontrunners in the race. The election was going to be decided by these two candidates only. This should be kept in mind when we build our models.

  Now datasets of county and state winners are created.
```{r, echo=F, message = F}
#Question 5

#county.winner
county.winner <- election.raw %>% 
  group_by(county) %>% 
  mutate(total=sum(votes)) %>%
  mutate(pct=votes/total) %>%
  top_n(n=1, wt=pct)
head(county.winner)

#state.winner
state.winner <- election.state %>%
  group_by(state) %>%
  mutate(total=sum(votes)) %>%
  mutate(pct=votes/total) %>%
  top_n(n=1, wt=pct)
head(state.winner)
```
  
  We can see Donald Trump seems to be the winner for most of the counties. Most of these counties are lower in population compared to counties in which Biden won, oweing to Trump's rural support compared to Biden's urban support.
  
# Visualization {.tabset}

##Generic State Map

```{r, cache = TRUE, echo=F, message = F}
#State-level map
states <- map_data("state")

ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group),
               color = "white") + 
  coord_fixed(1.3) +
  guides(fill=FALSE)  # color legend is unnecessary and takes too long


```
## Generic County Map
```{r,echo=F,message=F}
#Question 6: Use similar code to create county-level map
counties = map_data('county')

ggplot(data=counties)+
  geom_polygon(aes(x=long, y=lat, fill=region, group=group), color='white') +
  coord_fixed(1.3) +
  guides(fill=F)
```

## State Winners
```{r, echo=F, message = F}
#Question 7
state.winner$state <- tolower(state.winner$state)
states <- states %>% rename(state = region)

states.join <- left_join(states, state.winner)

ggplot(data=states.join)+
  geom_polygon(aes(x=long, y=lat, fill=candidate, group=group), color='white') +
  coord_fixed(1.3) +
  scale_fill_manual(values=c('orangered','steelblue')) + theme(legend.position='none')
```
 
  Using the map_data function from ggplot 2, a simple map of state winners is generated.

## County Winners
```{r,echo=F,message=F}
#Question 9
#Perform left_join
county.winner$county <- tolower(county.winner$county)
counties <- counties %>% rename(county = subregion)
counts.join <- left_join(counties,county.winner) 

ggplot(data=counts.join)+
  geom_polygon(aes(x=long, y=lat, fill=candidate, group=group), color='white') +
  coord_fixed(1.3) +
  scale_fill_manual(values=c('orangered','steelblue')) + theme(legend.position='none')
```
  
  Similar from the last map, county winners are also generated. Some of the data is missing for our county map, and is represented by the color grey. Particularly, Louisiana has a lot of missing counties.
  
## Californian County Winners
```{r echo=F, message=FALSE}
#Question 8

 

#Get California data
ca.counties <- filter(counts.join, region=='california')

ggplot(data=ca.counties)+
  geom_polygon(aes(x=long, y=lat, fill=candidate, group=group), color='white') +
  coord_fixed(1.3) +
  scale_fill_manual(values=c('orangered','steelblue')) + theme(legend.position='none')
#added custom colors for aesthetic
```
  

# Cleaning Up
```{r, echo=F,message=F}
#Question 10
#Filter out rows with missing values
census.clean <- census %>% na.omit()

#Convert {Men, Employed, VotingAgeCitizen} attributes to percentages
census.clean <- census.clean %>% 
  mutate(Men = Men/TotalPop*100) %>%
  mutate(Women=Women/TotalPop*100) %>%
  mutate(Employed = 100-Unemployment) %>%
  mutate(VotingAgeCitizen = (VotingAgeCitizen/TotalPop)*100)
  
#Compute Minority attribute
census.clean <- census.clean %>%
  mutate(Minority = Hispanic+Black+Native+Asian+Pacific)

census.clean <- select(census.clean, -c('Hispanic','Black','Native','Asian','Pacific','IncomeErr','IncomePerCap','IncomePerCapErr','Walk','PublicWork','Construction'))

#Print out first 5 rows
census.clean[1:5,]

```
  Now we need to clean the census data. Men, women, Employed, and VotingAgecitizen parameters were converted into percentages. Several parameters corresponding to the percentages of non-white races in each county were merged into a single Minority parameter. Several parameters were also removed.
  Some parameters were also removed due to perfect collinearity. These were Women (with Men) and Unemployment (with Employed)

  
# Dimensionality Reduction

```{r, echo=F}
#Question 11

census.clean.id <- select(census.clean,-c(State,County))
#Check if need to scale
summary( census.clean.id )
apply(census.clean.id, 2, var)
```

  Some percentages (such as Carpool, Transit, OtherTransp, WorkAtHome, etc) are much smaller on average compared to the averages of other percentage variables. There variances also differ significantly compared to each other. 
  
  If we didn't scale the variables before running PCA, then the PCs would be driven with the variables with the greatest mean and variance. While most parameters are already scaled for us (namely, the percentage parameters) we still need to consider TotalPop and Income
  
  Here's a list of the absolute values of the features for the First Component:
```{r, echo=F}
#Question 11

pr.out = prcomp(census.clean.id, scale=TRUE, center = TRUE)
#First two PCs
pcs.1_2 = pr.out$x[,1:2]
#Three features w/ largest absolute values for PC1.
pr.out$rotation[,1] %>% abs() %>% sort(decreasing=TRUE)
#Which features have opposite signs
```

  Poverty, ChildPoverty, and Employed have the largest absolute values of PC1. They have the highest correlations with the first component.
```{r, echo=F}
#Question 11

#Which features have opposite signs
pr.out$rotation[,1] %>% sort(decreasing=TRUE)
```

  The following features have opposite signs:
    Transit
    VotingAgeCitizen
    PrivateWork
    Men
    FamilyWork
    SelfEmployed
    Professional
    WorkAtHome
    Income
    White
    Employed
  
  This means they are all negatively correlated with the first principal component


```{r, echo=FALSE}
#Question 12

#Calculate PVE & Cumm PVE
pve = (pr.out$sdev^2)/sum( (pr.out$sdev^2) )
pve.cumm <- cumsum(pve)

#Plot PVE
plot(pve, xlab='PC', ylab='PVE', type='b', main='Proportion of Variance Explained(PVE) by each Principal Component(PC)')

#Plot Cumm PVE
plot(cumsum(pve), xlab='Principal Component', ylab='Cummulative PVE', ylim=c(0,1), type='b', main="Cummulative PVE by each Principal Component")
abline(v=which(pve.cumm >=.9)[1], col='red', lwd=3, lty=1)

```
  From the plot of the Cumulative PVE, we notice that the minimum number of PC's needed to capture 90% of the variance is 13 from a maximum possibility of 24.

# Clustering
```{r, echo=FALSE}
#Question 13

#calculate the euclidean distance matrix
cens.dist <- dist(census.clean.id)

#Run hierarchical clustering using complete linkage
cens.hclust <- hclust(cens.dist)

#Dataset of clustering information
census.clustinfo <- census.clean
census.clustinfo['cluster_10_norm'] <- cutree(cens.hclust, 10)
```
  
  Normally I would use a dendrogram to intrepret the clusters, but in this case it will just not be legible given the thousands of observations. If we want to observe how the clustering algorithm performed, we use the cutree function and put the clustered groups into a dataset with our information.
  
```{r,echo=F,include=F}
#Get ID of Santa Barbara County to for comparisons
SB.id = census.clean %>% filter(County == 'Santa Barbara') %>% select(CountyId)
SB.id=SB.id[[1,1]]
SB.id
```

```{r,echo=F,include=F}
#Clustering with normal parameters, 2 CLUSTERS

#Add new column to our copy of clustering info
census.clustinfo['cluster_2_norm'] <- cutree(cens.hclust, 2)
```

```{r,echo=F,include=F}
#Rerun clustering, replacing original features with pc1 & pc2

#calculate the euclidean distance matrix
pcs.dist = dist(pcs.1_2)

#Run hierarchical clustering using complete linkage
pcs.hclust <- hclust(pcs.dist)

#Plot the dendograms
dend.pc.1 = as.dendrogram(pcs.hclust)
dend.pc.1 = color_branches(dend.pc.1, k=10,groupLabels=T)
plot(dend.pc.1)
#Add new column to our copy of clustering info
census.clustinfo['cluster_10_pc'] <- cutree(pcs.hclust, 10)
```


```{r,echo=F,include=F}
#Rerun clustering, replacing original features with pc1 & pc2 USING 2 clusters

#Plot the dendograms
dend.pc.2 = as.dendrogram(pcs.hclust)
dend.pc.2 = color_branches(dend.pc.2, k=2,groupLabels=T)
plot(dend.pc.2)
#Add new column to our copy of clustering info
census.clustinfo['cluster_2_pc'] <- cutree(pcs.hclust, 2)
```

## Clustering Observations
```{r,echo=F,include=F}

census.clustinfo[census.clustinfo$CountyId == '6083',]
```
```{r}
#Second cluster, 10 total, normal parameters
census.clustinfo[census.clustinfo$cluster_10_norm == '2',]
```
```{r}
#first cluster, 2 total, normal parameters
census.clustinfo[census.clustinfo$cluster_2_norm == '1',]
```

```{r,echo=F,include=F}
#Second cluster, normal parameters
census.clustinfo[census.clustinfo$cluster_10_pc == '2',]
```

```{r,echo=F,include=F}
#First Cluster, normal parameters
census.clustinfo[census.clustinfo$cluster_2_pc == '1',]
```

# Classification

  We are now working with supervised learning tasks. Our most important question will be:
    Can we use census information in a county to predict the winner in that county?
  
## Getting the Data prepared
  We first need to combine county.winner and census.clean. 
```{r,message=F}
# we move all state and county names into lower-case
tmpwinner <- county.winner %>% ungroup %>%
  mutate_at(vars(state, county), tolower)

# we further remove suffixes of "county" and "parish"
tmpcensus <- census.clean %>% mutate_at(vars(State, County), tolower) %>%
  mutate(County = gsub(" county|  parish", "", County)) 

# we join the two datasets
election.cl <- tmpwinner %>%
  left_join(tmpcensus, by = c("state"="State", "county"="County")) %>% 
  na.omit

# drop levels of county winners if you haven't done so in previous parts
election.cl$candidate <- droplevels(election.cl$candidate)


## save meta information
election.meta <- election.cl %>% select(c(county, party, CountyId, state, votes, pct, total))

## save predictors and class labels
election.cl = election.cl %>% select(-c(county, party, CountyId, state, votes, pct, total))

#Question 14: Why do we need to exclude the predictor party from election.cl?
```

  The predictor party is excluded from the our classification data due to it's perfect collinearity with the response.
  
  We'll now split election.cl into a training and test data-set.
```{r}
set.seed(10) 
n <- nrow(election.cl)
idx.tr <- sample.int(n, 0.8*n) 
election.tr <- election.cl[idx.tr, ]
election.te <- election.cl[-idx.tr, ]
```

  And we'll perform 10-fold Cross Validation.
```{r}
set.seed(20) 
nfold <- 10
folds <- sample(cut(1:nrow(election.tr), breaks=nfold, labels=FALSE))
```

  Later on, we'll also need the following error rate function. The object records will be used to record the performance of our different methods.
```{r}
calc_error_rate = function(predicted.value, true.value){
  return(mean(true.value!=predicted.value))
}
records = matrix(NA, nrow=3, ncol=2)
colnames(records) = c("train.error","test.error")
rownames(records) = c("tree","logistic","lasso")
```

## Classification Models

### Decision Trees


```{r, echo=FALSE, fig.cap="Decision Tree before pruning"}
# Question 15

#Before Pruning
elect.tree <- tree(candidate~., data=election.tr)
draw.tree(elect.tree,nodeinfo=T,cex=.5,size=1.5)
```
  
```{r, echo=F}
# Question 15

#Get Error Rates of pre-pruned tree
elect.tree.pred <- predict(elect.tree, election.te, type='class')
paste("Test Error Rate:", calc_error_rate(elect.tree$y, election.te$candidate))

elect.tree.train_pred <- predict(elect.tree, election.tr, type='class')
paste("Train Error Rate:", calc_error_rate(elect.tree$y, election.tr$candidate))
```
  The Pre-pruned decision Tree splits the data at Transit first, and has a 33.45% test error rate with a 0% train error rate. The high test error rate is likely due to overfitting, since we're getting a train error rate of 0. Pruning the tree will help with our accuracy, and reduce overfitting. 
  We first need to determine the best size of the tree.
```{r,echo=F}
set.seed(20)
cv = cv.tree(elect.tree, FUN=prune.misclass, rand=folds)
best.cv = min(cv$size[cv$dev == min(cv$dev)])
best.cv
```

  The tree with 7 terminal nodes generates the lower error rate, so we will use that for pruning.
```{r,echo=F}
elect.pt <- prune.misclass ( elect.tree, best=best.cv)

draw.tree(elect.pt, cex=.7) +
  title('Pruned Tree w/ size=7')
```
  
  From the Pruned Tree shown above (and the Pre-Pruned Tree beforehand), "Transit" is split first. This may seem odd, but it likely has to do with its correlation with TotalPop. Public transit is much more common in cities, and cities contribute to high population in a county. From the graph, it appears that cities with a greater percentage of white people tend to have voted for Donald Trump, whereas cities with a lesser percentage of white people, meaning a higher percentage of minorities, voted for Joe Biden. Donald Trump had greatest success in counties with a lower percentage of transit use and higher percentage of white people. This would be rural counties with smaller cities. Rural, white voters overwhelmingly voted for Donald Trump, as we've seen from our visualizations and here.
  
  
  Now we can predict on the test set and calculate our error rate and record them in our table
```{r,echo=F}
#predict on test/train sets
elect.pt.predte = predict(elect.pt, election.te, type='class')
elect.pt.predtr = predict(elect.pt, election.tr, type='class')
#calculate test/train error rate
tree.err.te <- calc_error_rate(elect.pt.predte, election.te$candidate)
tree.err.tr <- calc_error_rate(elect.pt.predtr, election.tr$candidate)

records[1,1] <- tree.err.tr
records[1,2] <- tree.err.te
records
```
  We generate a training error rate of ~9% and a test error rate of ~13% through decision tree classification. This is a fairly good number, but we may yield better results through either the logistic or lasso methods.

### Logistic Regression

_"Well I believe I'll vote for a third party candidate!"_
_"Go ahead, throw your vote away! Ahahaha!"_
While the United States is not de jure a two party system, it is in practice. Fortunately, this allows us to create a logistic regression classification model. 

```{r,echo=F}
elect.glm <- glm(candidate~.,data=election.tr, family=binomial)
summary(elect.glm)
```
  Professional, VotingAgeCitizen and Service appear to have the greatest positive impact on the response. White, Drive, Carpool, and FamilyWork had the greatest negative impact. Our decision Tree had roughly the same variables as our logistic model.
 
  Before we can make predictions, we need to calculate the best threshold value to use. The "best" value would be the minimum euclidean distance from (FPR, FNR) to (0,0). Basically, we want the value that yields the lower FPR and FNR. We generate FPR and FNR versus threshold values, giving us a rough idea on what we can expect the threshold to be. The cutoff is found to be .2778, so we generate the smallest Euclidean distance. This will be our threshold. Now we can compute the error rates and record them.

```{r, message=FALSE, warning=FALSE, echo=F}
#predict on train
elect.glm.predtr <- predict(elect.glm,election.tr,type='response')
pred = prediction(elect.glm.predtr,election.tr$candidate)

#Calculate FPR, Cutoff and FNR
fpr = performance(pred,'fpr')@y.values[[1]]
cutoff = performance(pred,'fpr')@x.values[[1]]
fnr = performance(pred,'fnr')@y.values[[1]]

matplot(cutoff,cbind(fpr,fnr), type='l',lwd=2,xlab='Threshold',ylab='Error Rate')
legend(0.3,1,legend=c("False Positive Rate","False Negative Rate"), col=c(1,2),lty=c(1,2))
```
```{r, message=FALSE, warning=FALSE, echo=F}
rate = as.data.frame(cbind(Cutoff=cutoff,FPR=fpr,FNR=fnr))
rate$distance = sqrt((rate[,2]^2+rate[,3]^2))
threshold = rate$Cutoff[which.min(rate$distance)]

filter(rate, distance==min(distance))
```
```{r message=FALSE, warning=FALSE,echo=FALSE}
set.seed(20)

#make predictions
glm.pred.tr <- predict(elect.glm, election.tr,type='response')
glm.pred.te <- predict(elect.glm, election.te, type='response')

#convert predictions to classes
glm.pred.tr.class <- ifelse(glm.pred.tr>threshold,"Joe Biden", "Donald Trump")
glm.pred.te.class <- ifelse(glm.pred.te>threshold,"Joe Biden", "Donald Trump")

#Input into Records
records[2,1] <- calc_error_rate(glm.pred.tr.class, election.tr$candidate)
records[2,2] <- calc_error_rate(glm.pred.te.class, election.te$candidate)

records
```
  
Compared to the Tree Model, the training error rate is relative the same at 8%, but the test error rate is only 9.5%, a significant improvement. Finally, we should compare these results to a lasso model.

### Lasso Regression

We'll use the cv.glmnet function to perform a 10-fold cross validation and choose the best regularization parameter for the logistic regression with the LASSO penalty. 
```{r, echo=F}
x <- as.matrix( select(election.tr, -candidate) )
y <- as.matrix(select(election.tr,candidate))
y <-ifelse(y=="Joe Biden", 1,0)
set.seed(10)

cv.out = cv.glmnet(x, y, alpha=1, lambda=seq(1,50)*1e-4, family=binomial)

bestlam = cv.out$lambda.min
paste("best lambda value", bestlam)
```
  The optimal value of lambda is roughly 0.001. Now, we should check and see which coefficients are non-zero for the full dataset.
  
```{r,echo=F}
out = glmnet(x,y, alpha=0, family=binomial)

coefficients <- predict(out, type='coefficients', s=bestlam)
coefficients
```
  
  In the Lasso model, FamilyWork is the strongest coefficient, followed by OtherTransp, Transit, Professional, and Service. These all fall in line with what we've observed from our past models
  Finally, we compute the test and train error rates.  

```{r message=FALSE, warning=FALSE}
x.tr <- as.matrix(select(election.tr,-candidate))
y.tr <- as.matrix(select(election.tr,candidate))
y.tr <-ifelse(y.tr=="Joe Biden", 1,0)

#Lasso model, with best lambda
elect.lasso <- glmnet(x.tr, y.tr, alpha=1, lambda=bestlam, family=binomial)
#

#predict on train
lasso.tr <- predict(elect.lasso, newx=x.tr, type='response')
lasso.tr.class <- ifelse(lasso.tr >0.5, "Joe Biden", "Donald Trump")

#Predict on Test
x.te <- as.matrix(select(election.te, - candidate))
y.te <- as.matrix(select(election.te, candidate))
y.te <- ifelse(y.te == "Joe Biden", 1,0)
lasso.te <- predict(elect.lasso, newx = x.te, type='response')
lasso.te.class <- ifelse(lasso.te >0.5, "Joe Biden", "Donald Trump")

#Test/Train Error Rates
lasso.tr.err <- calc_error_rate(lasso.tr.class,election.tr$candidate)
lasso.te.err <- calc_error_rate(lasso.te.class, election.te$candidate)

paste("Lasso Model - Train Error Rate: ",lasso.tr.err)
paste("Lasso Model - Test Error Rate: ",lasso.te.err)

records[3,1] <- lasso.te.err
records[3,2] <- lasso.tr.err


```

We get a test error rate of 8.6% for our lasso model. That's pretty good! Let's check out our error rates for all of our models, now.
```{r}
records
```

  The Lasso model has the best test error rate of 7%. Our Logistic Regression model has a marginally worse test error rate, but slightly better train error rate. The decision trees model performed the worst in train and test error. Now, let's compute ROC curves using predictions on the test data for each of models.

```{r}
#Question 18

#Perform Predictions on testing data for each model
pred.tree <- prediction(as.numeric(elect.pt.predte), as.numeric(election.te$candidate))
pred.log <- prediction(as.numeric(glm.pred.te), as.numeric(election.te$candidate))
pred.lasso <- prediction(as.numeric(lasso.te), as.numeric(election.te$candidate))
#Compute Performances for each model
perf.tree <- performance(pred.tree, measure = 'tpr', x.measure = 'fpr')
perf.log <- performance(pred.log, measure='tpr', x.measure='fpr')
perf.lasso <- performance(pred.lasso, measure='tpr', x.measure='fpr')

#Plot ROC curves for each model

plot(perf.tree, col=2, lwd=3, main="ROC Curves")
plot(perf.log, add=TRUE, col=3, lwd=3)
plot(perf.lasso,add=TRUE,col=4,lwd=3)
abline(0,1)
legend('bottomright',legend=c("Dec Tree","Log Reg", "Lasso"), col=c(2,3,4), lty=3)
```
  
# Taking it Further (Classification)

  Since the Deicision Trees model performed the worst out of the three, let's see how a random forest model would perform.
## Random Forest
```{r, echo=F}
#Question 19
mtry = tuneRF(election.cl[,-1], election.cl$candidate)
elect.rf <- randomForest(candidate~.,data=election.tr,importance=TRUE,mtry = mtry )


elect.rf
```

```{r,echo=F}
importance(elect.rf)
```

   In random forest classification, generally, the number of variables randomly sampled at each split is the square root of the number of parameters. We use the tuneRF function to ensure that this holds true in this case. We can see that 1 variable was considered at each split for the lowest OOB error rate. This has to do with the mtry tuning.
   
  We can see that Transit was most important in terms of Model Accuracy(MeanDecreaseAccuracy), followed by Minority, White and TotalPop. Transit also was most important in terms of Gini Index (MeanDecreaseGini), again followed by Minority, White and TotalPop. It's worth noting that there exists a correlation between White and Minority due to the nature of their percentage values.
  
  This is similar to what we've observed in other models, especially in the logistic model which valued Transit, White, and TotalPop. 

## Boosting
  Let's make a boosted model as well. For a boosted model, we will set Joe Biden to be 1, and Donald Trump to be 0. We'll want 1000 bagged trees and set the interaction depth to be just 2. We'll leave the shrinkage parameter to its default of 0.001.
```{r,echo=F}
set.seed(27)
elect.boost <- gbm(ifelse(candidate=="Joe Biden", 1,0)~.,data=election.tr,
                   distribution='bernoulli',n.trees=1000,interaction.depth=2)
summary(elect.boost)
```
  Running the summary on the created model, we see that White is the most important variable in the model, followed by Income, Drive, and FamilyWork respectively. We can create a plot to visualize what the variables White and Income are doing to the response.
```{r,echo=F}
par(mfrow=c(1,2))
plot(elect.boost,i='White',type='response')
plot(elect.boost, i='Income', type='response')
```
  Remember, we set Donald Trump's value to be 0. It appears that the greater the percentage of White people make up a county, the county is more favorable in voting for Donald Trump. 
  As for income, counties appear to be more likely to vote for Joe Biden when average income is around 60k, 70k 80k. This is roughly middle, to upper-middle, class families.
  I want to note that the y-value observed in Income is extremely small, and I believe that has to do with the disproportional amount of favored-trumps that make up the Train data set. 
  
```{r,echo=F}
#Train Error Rate
rf.pred.tr <- predict(elect.rf, newdata = election.tr)
calc_error_rate(rf.pred.tr,election.tr$candidate)

#Test Error Rate
rf.pred.te <- predict(elect.rf, newdate=election.cl)
calc_error_rate(rf.pred.te,election.te$candidate)
```

# Regression

  With supervised classification done, we now move on to another task. The goal will now be:
    Can census information be used to predict the number of votes each candidate received per county?
  
  
```{r,echo=F,message=F}
# we move all state and county names into lower-case
tmpvotes <- election.raw %>% ungroup %>%
  mutate_at(vars(state, county), tolower)

# we move all state and county names into lower-case
# we further remove suffixes of "county" and "parish"
tmpcensus <- census.clean %>% mutate_at(vars(State, County), tolower) %>%
  mutate(County = gsub(" county|  parish", "", County))

# we join the two datasets
election.reg <- tmpvotes %>%
  left_join(tmpcensus, by = c("state"="State", "county"="County")) %>%
  na.omit


#ERROR GENERATED HERE
#Remove all observations of candidates that are not Joe Biden or Donald Trump
election.reg <- election.reg %>% filter(candidate==c('Joe Biden', 'Donald Trump'))

#Create dummy variable for if the observation is Joe Biden or Donald Trump
election.reg$biden <- ifelse(election.reg$candidate=='Joe Biden', 1,0)



# save predictors and class labels
election.reg = election.reg %>% select(-c(party))
#Remove all values where votes ==0
election.reg <- election.reg[election.reg$votes!=0,]
```
  NOTE: I am currently running into an error when filtering the dataframe of candidates that are not Joe Biden or Donald Trump. In previous versions of this Markdown, I hadn't ran into this error before. I need to examine more closely what is causing it and will do so at a further time.

  If we run our models as is, we are likely not going to get great results. As noted previously, Joe Biden and Donald Trump are the two candidates that obtained the vast majority of votes in the 2020 election. As such, our model should only focus on the vote totals of these candidates. Additionally, the candidates predictor is a nominal categorical variable. We need to mutate this variable into something numeric. So we use dummy variables to mutate the candidate variable into two variables that represent the characteristic. We NEED additional information for a regression model to work correctly. Not having a candidates variable will have each observation be identitcal in every value besides votes, which will likely lead to the same value being assigned twice per candidate for county observations.


## Modelling {.tabset}


```{r,echo=F,message=F,include=F}
#Records Table (For later use)
reg.records = matrix(NA, nrow=2, ncol=2)
colnames(reg.records) = c("train.MSE","test.MSE")
rownames(reg.records) = c("lasso","ridge")
```

```{r,echo=F,message=F,include=F}
#Splitting Data
n <- nrow(election.reg)
idx.tr <- sample.int(n, 0.8*n)
election.tr <- election.reg[idx.tr,]
election.te <- election.reg[-idx.tr,]
```


### LASSO MODELLING:

```{r,echo=F,message=F}
#Perform Cross Validation for hyperparameter tuning
#DO NOT LEAVE CANDIDATE FOR x, this will result in the matrix containing data of type 'character', not 'numeric' as desired
x <- as.matrix( select(election.tr, -c(county, votes, CountyId,candidate, state)) )
y <- election.tr$votes

set.seed(10)

cv.out = cv.glmnet(x, y, alpha=1, lambda=seq(1,50)*1e-4)

bestlam = cv.out$lambda.min
```

```{r,echo=F,message=F}
#Get train data
x.tr <- as.matrix(select(election.tr,-c(county, votes,CountyId,candidate, state)))
y.tr <- as.matrix(select(election.tr,votes))

#Lasso model, with best lambda
elect.lasso <- glmnet(x.tr, y.tr, alpha=1, lambda=bestlam)

#predict on train
lasso.pred.tr <- predict(elect.lasso, s= bestlam, newx=x.tr)


#Test Data w/ Predict on Test
x.te <- as.matrix(select(election.te, -c(county, votes,CountyId,candidate, state)))
y.te <- as.matrix(select(election.te, votes))
lasso.pred.te <- predict(elect.lasso,  s= bestlam, newx=x.te)

```

```{r,echo=F}
#Get Lasso Coefficients
predict(elect.lasso,type='coefficients',s=bestlam)
```
  It appears that there is a negative relationship between 'biden', "Minority", "Transit", and "White" with vote totals. This is somewhat expected, as most counties favored Donald Trump, though "White" does pique my interest.
```{r,echo=F}
paste("Percentage of counties that voted favored Trump: " ,mean(election.cl$candidate=='Donald Trump'))
```
  80% of the observations favored Donald Trump. Again, we know most of these counties had lower populations and were rural.
  
  For now, we'll just store Lasso results and move on.
```{r,echo=F}
#Train MSE
paste("Train MSE: ", round(mean((lasso.pred.tr-y.tr)^2,4)) )
reg.records[1,1] = round(mean((lasso.pred.tr-y.tr)^2,4))
paste("Test MSE: ",  round(mean((lasso.pred.te-y.te)^2,4)) )
reg.records[1,2] = round(mean((lasso.pred.te-y.te)^2,4))
```
### RIDGE REGRESSION

```{r,echo=F}
cor(x)
```
 
 There do exists some multicollinearities within the data, perhaps a ridge regression model may be appropriate.
 
```{r,echo=F,message=F}
#10-fold cv
lambda.list.ridge = 1000 * exp(seq(0,log(1e-5), length = 100))
ridge.cv = cv.glmnet(x.tr, y.tr, alpha=0, lambda=lambda.list.ridge)
#Get best lambda
bestlam = ridge.cv$lambda.min

#Ridge-Regression Model fitted to our training set

elect.ridge = glmnet(x.tr,y.tr, alpha=0,lambda=bestlam)

predict(elect.ridge,type='coefficients',s=bestlam)
```
  
  With our ridge model, it appears that PrivateWork and Men percentages have the greatest impact on the number of votes, followed by Women, Product, and Drive predictors.
  
```{r,echo=F,message=F}
#Ridge - Train MSE
ridge.pred.tr = predict(elect.ridge, s=bestlam, newx=x.tr)
paste("Train MSE: ", round(mean((ridge.pred.tr-y.tr)^2),4) )

reg.records[2,1] = round(mean((ridge.pred.tr-y.tr)^2),4)
#Ridge - Test MSE
ridge.pred.te = predict(elect.ridge, s=bestlam, newx=x.te)
paste("Test MSE: ", round(mean((ridge.pred.te-y.te  )^2),4) ) 

reg.records[2,2] = round(mean((ridge.pred.te-y.te)^2),4)
```
  
  Let's take a look at the results for the regularized regression models.
  
```{r,echo=F}
reg.records
```

  It doesn't seem like the ridge model faired better than our lasso approach. Let's stick with the lasso model to see how well it did in it's total predictions.
  
```{r,echo=F,message=F}
#create a single vector of lasso predictions
lasso.pred <- c(lasso.pred.tr,lasso.pred.te)

```

```{r,echo=F,message=F}
#remerge the train and test datasets, assign a column of predicted votes for our lasso predictions
election.reg <- rbind(election.tr,election.te)
election.reg['votes_pred'] <- lasso.pred
election.reg

#Similar to the initial analysis, create dataframes of predicted county and state values
election.reg.county <- election.reg
election.reg.state <- election.reg %>% group_by(state,candidate) %>% summarise(votes_pred=sum(votes_pred))

```


```{r,echo=F,message=F}
#county.winner (predicted)
pred.county.winner <- election.reg.county %>% 
  group_by(county) %>% 
  mutate(total=sum(votes_pred)) %>%
  mutate(pct=votes_pred/total) %>%
  top_n(n=1, wt=pct)
head(pred.county.winner)

#state.winner (predicted)
pred.state.winner <- election.reg.state %>%
  group_by(state) %>%
  mutate(total=sum(votes_pred)) %>%
  mutate(pct=votes_pred/total) %>%
  top_n(n=1, wt=pct)
head(pred.state.winner)
```
  
  The top dataset is the head of predicted county winners, the bottom are the states. 
```{r,echo=F,message=F,include=F}
#Checking full datasets
pred.county.winner

pred.state.winner
```

```{r,echo=F,message=F}
#Predicted Results
states.join <- left_join(states, pred.state.winner)

ggplot(data=states.join)+
  geom_polygon(aes(x=long, y=lat, fill=candidate, group=group), color='white') +
  coord_fixed(1.3) +
  scale_fill_manual(values=c('orangered','steelblue')) + theme(legend.position='none')
#added custom colors for aesthetic 
```

  So, it doesn't seem like the Lasso model did too well in accurately predicting election results. Each statewide election was fairly close but Donald Trump was the winner in each state, besides Rhode Island. Some values predicted are also running negative. My knowledge has run its limit, and at this time I am not aware on how to adjust my model to achieve better results. I will look into this in time. For now, I will use the knowledge I have in regression analysis to make my own model.

# Regression Analysis

## Model Building {.tabset}

### Stepside Regression
```{r,echo=F,include=F}
election.reg <- na.omit(election.reg)
x1 <- election.reg$TotalPop
x2 <- election.reg$Men

x4 <- election.reg$White
x5 <- election.reg$Minority
x6 <- election.reg$Poverty
x7 <- election.reg$ChildPoverty
x8 <- election.reg$biden

x10 <- election.reg$Employed

x12 <- election.reg$VotingAgeCitizen
x13 <- election.reg$Income
x14 <- election.reg$Professional
x15 <- election.reg$Service
x16 <- election.reg$Office
x17 <- election.reg$Production
x18 <- election.reg$Drive
x19 <- election.reg$Carpool
x20 <- election.reg$Transit
x21 <- election.reg$OtherTransp
x22 <- election.reg$WorkAtHome
x23 <- election.reg$MeanCommute
x24 <- election.reg$PrivateWork
x25 <- election.reg$SelfEmployed
x26 <- election.reg$FamilyWork


y <- election.reg$votes

#Obtain a model using stepwise regression

mod.0 = lm(y~1)
mod.U=lm(y~x1+x2+x4+x5+x6+x7+x8+x10+x12+x13+x14+x15+x16+x17+x18+x19+x20+x21+x22+x23+x24+x25+x26)

s.lm <- step(mod.0, scope = list(lower=mod.0, upper=mod.U))
```

```{r,echo=F}
summary(s.lm)
```
  Stewpwise regression found x1,x20,x14,x24,x16,x15,x23,x22 to be the best predictors for vote totals. These predictors, respectively, are:

    TotalPop - This should be expected, as there is a correlation between Totalpop and votes
    Transit - Similar to what we found in past models
    Professional
    PrivateWork
    Office
    Service
    MeanCommute
    WorkAtHome
    SelfEmployed

  83% of the variance is explained with the model. We do have a wide range of residuals which may be an issue. Furthermore, biden and trump predictors were not found to be statistically signifigant, which may cause trouble down the line with predictions. 

### Best Subsets  
```{r,echo=F}
library(leaps)
#Use best subsets regression and compare subsets with Mallow's Cp and adjusted R-squared, Compare with stepwise

mod = regsubsets(cbind(x1,x2,x4,x5,x6,x7,x8,x10,x12,x13,x14,x15,x16,x17,x18,x19,x20,x21,x22,x23,x24,x25,x26), y, nvmax = 26)
summary.mod <- summary(mod)
summary.mod$which
summary.mod$cp
summary.mod$adjr2
```
```{r,echo=F}
plot(summary.mod$bic,xlab='parameter',ylab='BIC')
```
```{r,echo=F}
plot(summary.mod$cp,xlab='parameter',ylab='Mallows Cp')
```
```{r,echo=F}
plot(summary.mod$adjr2,xlab='parameter',ylab='Mallows Cp')
```

  We want to favor a subset model with relatively low Cp/BIC valuess and relatively high aRSS. We can gather that the subset of 9 variables, the best of the 9-predictors models, is a good choice, these variables are also the same as found in the stepwise regression model, confirming results.

  Now we have a rough idea on what the model should look like.
```{r,echo=F}
reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)
```

Let's now check on the correlations between these variables
```{r,echo=F}
#Better Identify variables
votes <- election.reg$votes
TotalPop <- x1
Transit <- x20
Professional <- x14
PrivateWork <- x24
Office <- x16
Service <- x15
MeanCommute <- x23
WorkAtHome <- x22
SelfEmployed <- x25

#Scatterplot matrix
pairs(~votes+TotalPop+Professional+Service+Office+Transit+WorkAtHome+MeanCommute+PrivateWork+SelfEmployed)
```
```{r,echo=F}
#Correlation Matrix
subset.reg <- data.frame(votes,TotalPop,Professional,Service,Office,Transit,WorkAtHome,MeanCommute,PrivateWork,SelfEmployed)
cor(subset.reg)

```
  
  We can see Total Pop has a strong correlation with votes. Most other correlations are fairly weak. 
There's a fair positive correlation between WorkAthome and SelfEmployed, and a fair negative correlation between SelfEmployed and PrivateWork. F-tests can be run to see if interaction terms between these two pairs would be beneficial to the model.

### Interaction Terms
```{r,echo=F}
#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#Model for interaction between WorkAtHome and SelfEmployed
reg.interact.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+I(x22*x25))

#General Linear F-test
anova(reg.1,reg.interact.1)


#Model for interaction between WorkAtHome and SelfEmployed
reg.interact.2 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+I(x24*x25))

#General Linear F-test
anova(reg.1,reg.interact.2)
```
For the WorkAtHome*SelfEmployed interactionWe use our initial model, reg.1 as a null hypothesis, with our interaction model as the alternative. We obtain an F-statistic of 2.2974 with a p-value of .1293. So we fail to reject the null hypothesis. 

Now we check the interaction term between SelfEmployed and PrivateWork, and obtain an F-statistic of 1.3428 with a resulting p-value of .2476. Again, the null hypothesis isn't rejected.

Finally we should see if high order terms should be included in the model.

### Higher Order Terms
```{r,echo=F}
#Testing Second Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#TotalPop, second order
x1_2 <- x1**2
summary(lm(y~x1+x1_2))

#Professional, second order
x14_2 <- x14**2
summary(lm(y~x14+x14_2))

#Service, second order
x15_2 <- x15**2
summary(lm(y~x15+x15_2))

#Office, second order
x16_2 <- x16**2
summary(lm(y~x16+x16_2))

#Transit, second order
x20_2 <- x20**2
summary(lm(y~x20+x20_2))

#WorkAtHome, second order
x22_2 <- x22**2
summary(lm(y~x22+x22_2))

#MeanCommute, second order
x23_2 <- x23**2
summary(lm(y~x23+x23_2))

#PrivateWork, second order
x24_2 <- x24**2
summary(lm(y~x24+x24_2))

#SelfEmployed, second order
x25_2 <- x25**2
summary(lm(y~x25+x25_2))
```
All second order terms besides x16, Office, seem to need further testing. We put all signfigant second order terms into a new model and anova test to see if we should use the new model, labeled reg.2, or our initial model without second order terms, reg.1.

```{r,echo=F}
reg.2 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2)
anova(reg.1,reg.2)
```
The p-val generated from the test is less than 0.05, so the Null Hypothesis is rejected, we use our new model.

No we do further testing for signifigant third order terms.
```{r,echo=F}
##Testing third Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#TotalPop, third order
x1_3 <- x1**3
summary(lm(y~x1+x1_2+x1_3))

#Professional, third order
x14_3 <- x14**3
summary(lm(y~x14+x14_2+x14_3))

#Service, third order
x15_3 <- x15**3
summary(lm(y~x15+x15_2+x15_3))



#Transit, third order
x20_3 <- x20**3
summary(lm(y~x20+x20_2+x20_3))

#WorkAtHome, third order
x22_3 <- x22**3
summary(lm(y~x22+x22_2+x22_3))

#MeanCommute, third order
x23_3 <- x23**3
summary(lm(y~x23+x23_2+x23_3))

#PrivateWork, third order
x24_3 <- x24**3
summary(lm(y~x24+x24_2+x24_3))

#SelfEmployed, third order
x25_3 <- x25**3
summary(lm(y~x25+x25_2+x25_3))
```
Third order interactions for MeanCommute and SelfEmployed don't seem to be signfigant. Once we use an anova test to interpret model improvement.
```{r,echo=F}
reg.3 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3)
anova(reg.2,reg.3)
```
```{r,echo=F}
##Testing fourth Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#TotalPop, fourth order
x1_4 <- x1**4
summary(lm(y~x1+x1_2+x1_3+x1_4))

#Professional, fourth order
x14_4 <- x14**4
summary(lm(y~x14+x14_2+x14_3+x14_4))

#Service, fourth order
x15_4 <- x15**4
summary(lm(y~x15+x15_2+x15_3+x15_4))

#Transit, fourth order
x20_4 <- x20**4
summary(lm(y~x20+x20_2+x20_3+x20_4))

#WorkAtHome, fourth order
x22_4 <- x22**4
summary(lm(y~x22+x22_2+x22_3+x22_4))

#PrivateWork, fourth order
x24_4 <- x24**4
summary(lm(y~x24+x24_2+x24_3+x24_4))

```
Fourth Order Terms for TotalPop and Service (x1 & x15) are not signifigant. We once again run an anova test.
```{r,echo=F}
reg.4 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3+
              x14_4+x20_4+x22_4+x24_4)
anova(reg.3,reg.4)
```

```{r,echo=F}
##Testing fifth Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)



#Professional, fifth order
x14_5 <- x14**5
summary(lm(y~x14+x14_2+x14_3+x14_4+x14_5))

#Transit, fifth order
x20_5 <- x20**5
summary(lm(y~x20+x20_2+x20_3+x20_4+x20_5))

#WorkAtHome, fifth order
x22_5 <- x22**5
summary(lm(y~x22+x22_2+x22_3+x22_4+x22_5))

#PrivateWork, fifth order
x24_5 <- x24**5
summary(lm(y~x24+x24_2+x24_3+x24_4+x24_5))
```
Fifth order terms for Professional, WorkAtHome, and PrivateWork are not signifigant. 
```{r,echo=F}
reg.5 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3+
              x14_4+x20_4+x22_4+x24_4+
              x20_5)
anova(reg.4,reg.5)
```
Adding x20_5 to the model is a statistically signfigiant improvement, we continue adding higher order terms for x_20
```{r,echo=F}
##Testing sixth Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#Transit, sixth order
x20_6 <- x20**6
summary(lm(y~x20+x20_2+x20_3+x20_4+x20_5+x20_6))


```
```{r,echo=F}
reg.6 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3+
              x14_4+x20_4+x22_4+x24_4+
              x20_5+
              x20_6)
anova(reg.5,reg.6)
```
```{r,echo=F}
##Testing seventh Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#Transit, sixth order
x20_7 <- x20**7
summary(lm(y~x20+x20_2+x20_3+x20_4+x20_5+x20_6+x20_7))
```
```{r,echo=F}
reg.7 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3+
              x14_4+x20_4+x22_4+x24_4+
              x20_5+
              x20_6+
              x20_7)
anova(reg.6,reg.7)
```
```{r,echo=F}
##Testing eigth Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#Transit, eigth order
x20_8 <- x20**8
summary(lm(y~x20+x20_2+x20_3+x20_4+x20_5+x20_6+x20_7+x20_8))
```
```{r,echo=F}
reg.8 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3+
              x14_4+x20_4+x22_4+x24_4+
              x20_5+
              x20_6+
              x20_7+
              x20_8)
anova(reg.7,reg.8)
```
```{r,echo=F}
##Testing ninth Order terms

#(Reference) Initial model
#reg.1 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25)

#Transit, ninth order
x20_9 <- x20**9
summary(lm(y~x20+x20_2+x20_3+x20_4+x20_5+x20_6+x20_7+x20_8+x20_9))
```
```{r,echo=F}
reg.9 <- lm(y~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3+
              x14_4+x20_4+x22_4+x24_4+
              x20_5+
              x20_6+
              x20_7+
              x20_8+
              x20_9)
anova(reg.8,reg.9)

```
We finally fail to reject the null hypothesis using anova testing, and thus have a final model.
```{r,echo=F}
reg.full <- reg.8
```

## Residual Analysis
  
  Before any estimations, we need to ensure LINE requirements so that our assumptions are met.
 
```{r,echo=F}
#Get residuals and fitted values, create residuals v. fit plot
res <- resid(reg.full)
fit <- fitted(reg.full)

plot(x=fit,y=res, xlab='fitted values', ylab='residual values', main='Residual V. Fit Plot')+
  abline(h=0,col='red')
```
  
  There is a clear nonconstant variances with our residuals, along with noticable outliers. We have a fanning effect, with our residuals being close to 0 for small x-values, but increasing with larger ones.
  
```{r,echo=F}
#qqnorm Plot
qqnorm(res)
qqline(res,col='blue')
```

The qqnorm plot indicates our residuals are not normally distributed. There's also clear outliers in the data that's likely affected the model.
Let's create a box-cox error plot to validate these claims. If our estimated value of lambda has 1 within it's 95% confidence interval, then we may assume normality. If not, we'll need to transform the data in some way to fix this.

```{r,echo=F,message=F}
library(MASS)
bcox <- boxcox(reg.full)
bcox
```
  
  We can also use a shapiro-wilks test. In the shapiro-wilks test, we assume for the null hypothesis that residuals are normally distributed. We obtain a p-value less than 0.05, thus rejecting the null hypothesis. The residuals are clearly not normally distributed.
  
```{r,echo=F}
shapiro.test(res)
```
  
  So it is true, our residuals are not normally distributed and as such LINE conditions are not met. We'll need to transform our response in some way to fulfill LINE conditions. Since 0 doesn't fall into the 95% confidence interval for the boxcox transform, we'll take the value of lambda we have and use a boxcox transformation on the votes parameter.
  
```{r,echo=F}
lambda <- bcox$x[which.max(bcox$y)]
paste("Boxcox Lambda: ",lambda)
```
```{r,echo=F}
reg.transform <- lm(y**lambda~x1+x20+x14+x24+x16+x15+x23+x22+x25+
              x1_2+x14_2+x15_2+x20_2+x22_2+x23_2+x24_2+x25_2+
              x1_3+x14_3+x15_3+x20_3+x22_3+x24_3+
              x14_4+x20_4+x22_4+x24_4+
              x20_5+
              x20_6+
              x20_7+
              x20_8)
```

```{r,echo=F,message=F}
#New residual v. fit plot
res <- resid(reg.transform)
fit <- fitted(reg.transform)

plot(x=fit,y=res, xlab='fitted values', ylab='residual values', main='Residual V. Fit Plot (Box-Cox Transform)')+
  abline(h=0,col='blue')
```
  The residuals seem to be much more constant than before, but we still have a cone shape. 
```{r,echo=F}
#qqnorm Plot
qqnorm(res)
qqline(res,col='blue')
```
  Our data appears much more normally distributed, but there's still an apperent left skew.

```{r,echo=F}
shapiro.test(res)
```
  The shapiro wilks test still rejects the null hyothesis. At this point, my knowledge on the subject has ran its course. There's a case to consider outliers and high leverage data, there are noticable outliers and possible high leverage data that we can observe from our plots and would likely find in further analysis. However, there is no reason to assume the given data is incorrect, and so they should not be excluded from the data. 

  I'm going to check the models predictions and see how it performed.
```{r,echo=F}
#Get our regression data, minus votes
election.reg.data <- data.frame(TotalPop,Professional,Service,Office,Transit,WorkAtHome,MeanCommute,PrivateWork,SelfEmployed)
#predict the data and add as a column to our original data frame.
reg.pred <- predict(reg.transform,election.reg.data)


```
```{r}
election.reg$reg.pred <- reg.pred**(1/.22)
election.reg
```

```{r,echo=F, message = F}
#Similar to the initial analysis and regularized regression models, create dataframes of predicted county and state values
election.reg.county <- election.reg
election.reg.state <- election.reg %>% group_by(state,candidate) %>% summarise(reg.pred=sum(reg.pred))

#county.winner (predicted)
pred.county.winner <- election.reg.county %>% 
  group_by(county) %>% 
  mutate(total=sum(reg.pred)) %>%
  mutate(pct=reg.pred/total) %>%
  top_n(n=1, wt=pct)
head(pred.county.winner)

#state.winner (predicted)
pred.state.winner <- election.reg.state %>%
  group_by(state) %>%
  mutate(total=sum(reg.pred)) %>%
  mutate(pct=reg.pred/total) %>%
  top_n(n=1, wt=pct)
head(pred.state.winner)
```
  The header of county winners is printed above, and the header of state winners is printed below. The results are not promising.
```{r,echo=F,include=F}
pred.county.winner
pred.state.winner
```
  
```{r,echo=F}
#Predicted Results
states.join <- left_join(states, pred.state.winner)

ggplot(data=states.join)+
  geom_polygon(aes(x=long, y=lat, fill=candidate, group=group), color='white') +
  coord_fixed(1.3) +
  scale_fill_manual(values=c('orangered','steelblue')) + theme(legend.position='none')
#added custom colors for aesthetic 
```
  There's a tie in virtualy state, likely the dummy variable was not included in the final model. There wasn't enough information to for more accuracte estimations.


  If further work in this analysis were to be done, I would seriously consider the inclusion of the Biden dumby variable regardless of statistical signfigance. That seems to be the only way to make meaning of the difference of votes for each county.

  My knowledge of model building is admitedly limited to my knowledge from my class on regression analysis, and I'd need to further study to make sure my process is correct. As far as I know, it is, but there's always room for improvement. As it stands, this model doesn't have enough information to make meaningful predictions on election results, though it does seem to perform well in predicting votes per county.



# Final Thoughts

  My work on classification seems to have done a better job in making accurate predictions towards election results. For that I can say it's relatively small error rate has me satisfied. The Lasso classification model performed the best, but the logistic regression and random forest models also performed well. The decision trees model performed the worst of the classification models.
  
  For Regularized Regression, the Lasso model performed best, but still not well. As it currently stands my regression model did not perform too well, either. It would be fair to say that that census data does a decent enough job at predicting county winners, but poorly at predicting the number of votes from each county.

  The Transit predictor was the most important of all census data, likely because of it's relation to urban populations and demecratic majorities. Transit also typically had the greatest coefficients in the regression models.Higher rates of Transit and Population typically meant the county would favor Biden, according to the decision trees model. Higher rates of White typically favored Trump. Other variables of interest for classification included FamilyWork, VotingAgeCitizen, Service, and Minority.
  
  

  
  